---
title: "Monarch Watch"
author: "Name"
format: html
editor: visual
message: false
warning: false
---

```{r}
#| include: false
#install.packages("googlesheets4")
library(googlesheets4)
library(tidyverse)
library(janitor)
library(tidygeocoder)
library(sf)
library(leaflet)
library(knitr)
```

Every year Monarch Butterflies fly south for the winter. They start in higher latitudes and migrate south to their wintering areas in Mexico.

[Monarch Watch](https://monarchwatch.org/blog/) is a citizen science initiative housed in the University of Kansas. Every year volunteers tag these butterflies and release them so their progress in their southern migration can be mapped. If someone finds a monarch with a tag they [can report it.](https://docs.google.com/forms/d/e/1FAIpQLSetTI4HYcyb1GfIyo-U5rAuVVezvOrwicz2ztIt244Vh7S4VA/viewform) We examine only the Eastern Monarch's trip south and create a map showing the areas the butterflies are sighted in the US.

```{r}
#| include: false
# US and Canada Data
US_CAN_Recovery_2022 <- read_sheet("https://docs.google.com/spreadsheets/d/14ONbP-0rgvVz-DR0MWYkkPm0BzyFBHMiNCA2F5I8Yks/edit#gid=1710298744")

#Mexico Recovery Data
Pre_2023_Mexico_Recovery <- read_sheet("https://docs.google.com/spreadsheets/d/1UdJfooBJrm0Y1zlpwIhGz7ToZfP9h8OeucJbXrWwZEY/edit#gid=1853245517")

# If you cannot get acces to the above data run these two lines in the console. Otherwise ignore this message
# googlesheets4::gs4_deauth()
# googlesheets4::gs4_auth()

# The code below takes both of these dataframes and makes them into a list. 

data_frame_list <- list(US_CAN_Recovery_2022, Pre_2023_Mexico_Recovery)
```

```{r}
#| include: false
# We need to clean the data we downloaded from google sheets. We can use clean_names() to do this. Since there are two dataframes we can clean them both at once. Use map() to do so. 

# Iterate over your function to clean the names and store the three dataframes as a list. 
butterfly_dataframes <- map(
  .x = data_frame_list ,#add the name of the list here,
  .f = clean_names #add the name of the function here. 
)

# At the moment butterfly_dataframes is unnamed which is inconvient so below I'm naming all the dataframes in the list.  

names(butterfly_dataframes) <-   c( "US_CAN_Recovery_2022", "Pre_2023_Mexico_Recovery")
```

```{r}
#| include: false
USA_map <-  maps::map("state", plot = FALSE, fill = TRUE) |> 
  st_as_sf(coords = c("x", "y"), crs = 4326)
```

## A Chrolopleth map of the US.

```{r}
#| include: false
#butterfly_location_data <- butterfly_dataframes$US_CAN_Recovery_2022 |>
 # select(tag_code,city_location,state_province,country, date)|>
 # geocode(state = state_province, country = country)

#save(butterfly_location_data, file = "butterfly_location_data_city.Rdata")
# 
load("butterfly_location_data_city.Rdata")

```

```{r}
#| include: false
# Wrangle your data
butterfly_location_summary <- butterfly_location_data |>
   group_by(state_province) |>
   summarise(
          total = nrow(x=butterfly_location_data),
          sighting= n(), #total sighting per state here
          total_prop= sighting/total)# total proportions
          


# Here is a df with state abreviation and names
states<- as.data.frame(cbind(abrev = state.abb, lower_state = tolower(state.name), upper_state = state.name))

#First join
full_joined_df <- 
  full_join(states, butterfly_location_summary, by= c("abrev"="state_province"))


#Second join  Inner join to get the location data. Call the resulting dataframe USA_map_joined.
USA_map_joined <-  inner_join(USA_map, full_joined_df, by= c("ID"="lower_state"))
  

```

```{r}
#| echo: false
#| label: fig-1
#| fig-cap: "The map shows butterfly sightings in each US state"
# Make map here.

  # The code below set a color palette and changes the background color. Orange colors for butterflys and light blue for the background the make the points pop. I also turned off the grid lines. 

USA_map |> ggplot()+
  geom_sf()+
  geom_sf(data= USA_map_joined, aes(fill= sighting, color = "black"))+
  scale_fill_distiller(palette= "YlOrRd", direction= 1)+
  labs(title= "Butterfly Sightings by State")+
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white",
                                colour = "black",
                                size = 0.5),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text = element_blank())
  

```
