---
title: "Querying a MySQL database"
format:  html
  
execute: 
  echo: false
---

```{r global_options, include=FALSE}
#| include: false
posted <- Sys.time() > "2020-11-25 11:00:00 EDT"
```

Using SQL queries on the school's server, I have compiled data related to the filmographies of actresses Andie MacDowell and Nicole Kidman.

```{r, message=FALSE}
#| include: false
#| 
library(tidyverse)
library(RMySQL)
db <- dbConnect(
  MySQL(),
  host = "scidb.smith.edu",
  user = "sds192",
  password = "DSismfc@S",
  dbname = "imdb"
  )
knitr::opts_chunk$set(connection= 'db', max.print = 5)
```


```{r}
#| include: false

class(db)
```


```{sql, connection='db' , include=FALSE}

USE imdb;
```
 
For Andie MacDowell, I've created a table detailing her complete filmography, including the title of each movie, its production year, and the character she portrayed, arranged chronologically.


```{sql, connection='db', echo= FALSE}

SELECT t.title, t.production_year, cn.name
FROM cast_info ci
JOIN  char_name cn  ON cn.id = ci.person_role_id
JOIN   title t ON t.id = ci.movie_id
WHERE ci.person_id = 3479179
  AND t.kind_id = 1
ORDER BY production_year ;
```



In the following table, I've extracted data from a webpage listing Nicole Kidman as the most prolific actress, specifically capturing the names and production years of all the movies she's been in, ordered by production year in descending order.

```{sql,connection='db', echo=FALSE}
SELECT t.title, t.production_year, cn.name
FROM cast_info ci
JOIN  char_name cn  ON cn.id = ci.person_role_id
JOIN   title t ON t.id = ci.movie_id
WHERE ci.person_id = 3353717
  AND t.kind_id = 1
ORDER BY production_year DESC;
```
